Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-05-10 ms/response_letter.Rmd

> AE comments: Thanks for the submission. I agree with the reviewers that the paper has merit and will be of interest to many researchers because LMA is a widely used trait. Both reviewers have some feedback for you to consider. Reviewer 1's suggestions would require extensive revisions to address. Reviewer 2 has suggested a modification to the title. I think the existing title is fine if you prefer to keep that but would ask you to carefully consider all other comments carefully.

> Reviewer #1: Summary of Manuscript:
> This manuscript compares two methods of measuring LMA (leaf mass per area), an ecologically important functional trait in plants. It is a carefully executed and useful piece of work, in that LMA is a very widely measured plant trait and many readers will be interested in identifying the best methods to use when measuring LMA in their system, as well as how to compare their own measurements to other published estimates. However, I believe the paper needs modification; I include recommendations, including three major concerns, below:

We thank the reviewer for the encouraging comments and helpful critiques.

@@

> 1. Framing: I think the paper misses an opportunity to provide a bit more context for why the study is important. The authors could provide more concrete examples of why standardizing LMA techniques across studies matters: e.g. what is the ecological significance of an LMA measurement difference between the two methods?  This could also be set up in the introduction by providing more context for LMA as a functional trait. For example, by citing the TRY database, the authors could provide readers with the importance and ubiquity of LMA relative to other plant traits (e.g. of all the studies of plant traits in TRY, how many report LMA?)

- record and requests

> 2. A crucial thing for readers to know is how to "convert" one method's estimate of LMA to another: the authors could report an equation (or a few, accounting for some additional covariates such as leaf area and leaf thickness, which this paper identifies as important modifiers of the relationship) that scholars could use in the future if trying to compare LMA estimates collected using different methods. In other words, taking the eq from Appendix 2 and modifying it given the results of this study.

Thank you.
This is an important thing to clarify.
We now use a linear model that used leaf area and leaf thickness as covariates but we suggest OLS models without covariates for the conversion of LMA because leaf thickness and leaf area affect not only the mean but also the variance in the relationship between the leaf disc and whole-leaf estimates, and that is why R2 values changed among the four groups (thin vs. think and small vs. large leaves) and it is not to practical to convert LMA using leaf area or leaf thickness.
We also present a practical guideline for conversion of LMA.

@@

> Line-specific Comments:

> Line 1: Suggest adding "- evidence from 334 woody species" in the title.

> Lines 26 and 51: Change "the key trait" to "an important functional trait".

Thank you.
We have updated the line 26 but key the line 51 as "the key trait" because xxx

### 2022-05-10 ms/LMA_method.Rmd
update ms

**PREMISE:**
Leaf mass per area (LMA), which is the key trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

@@

Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area thus reflecting a cost for light interception, which is the key trait of the leaf economic spectrum (LES) [@Wright2004a; @Poorter2009; @Onoda2017].
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover [@Wright2004a].
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns [@Garnier1994; @Wright2004a; @Reich2014; @Diaz2016].
The LMA is also frequently used in plant growth analysis [@Evans1972; @Poorter2014; @Falster2018] because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR $\times$  LMR $\times$ LMA^-1^).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species [i.e., 'soft' trait, @Diaz2004].
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago [@Hanson1917].

LMA values can be determined by either
1) a whole leaf including or excluding petioles, and
2) a leaf disc excluding all major veins or petioles [@Perez-Harguindeguy2013; @Kattge2020].
Although some studies do not make clear what protocol was followed, these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA [@John2017].
Since larger leaves tend to invest more of their mass into dense midribs and petiole for support [@Niinemets2006; @Niinemets2007], and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; [@Grubb1986]), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
If leaf disc LMA largely and consistently underestimates whole-leaf LMA, some calibrations may be required to combine or compare those different estimates [@Kraft2008; @Onoda2011].
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., low *R^2^* values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database [@Wright2010], is available in the TRY [@Kattge2020] that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.

>>>>

**PREMISE:**
Leaf mass per area (LMA), which is an important functional trait in leaf economic spectrum and plant growth analysis, is measured from leaf discs or whole leaves.
These differences in the measurement methods may lead to large differences in the estimates of LMA values.

@@

The primary function of leaves is to return photosynthetic revenue on the investment that has been made in constructing the leaf [@Westoby2013], which is inherently economic in nature [i.e., leaf economic spectrum, @Wright2004a].
Leaf mass per area (LMA), determined by lamina thickness and leaf tissue density (LD), describes how much biomass is invested into given photosynthetic leaf area, which is the key trait of the leaf economic spectrum [@Wright2004a; @Poorter2009; @Onoda2017].
Generally, resource-acquisitive species (fast-growing species) tend to have low LMA values, showing high photosynthesis, high nutrients, and often have fast leaf turnover [@Wright2004a].
In contrast, resource-conservative species (slow-growing species) often have higher LMA values with the opposite patterns [@Garnier1994; @Wright2004a; @Reich2014; @Diaz2016].
The LMA is also frequently used in plant growth analysis [@Evans1972; @Poorter2014; @Falster2018] because relative growth rates (RGR) can be decomposed into the product of net assimilation rate (NAR), leaf mass ratio (LMR) and LMA (i.e., RGR = NAR $\times$  LMR $\times$ LMA^-1^).
Another appealing feature of LMA is that it is relatively easy to measure large numbers of species [i.e., 'soft' trait, @Diaz2004].
Therefore, LMA has been of interests to ecologists and widely used since the first report more than a century ago [@Hanson1917].
Actually, LMA was the most often requested leaf trait in the TRY plant data database (2,977 out of 7,330 requests), followed by leaf nitrogen contents per leaf dry mass [1,938 requests, @Kattge2020 [status October 1 2019]].

LMA values can be determined by either
1) a whole leaf including (88,490 observations and 7,068 species) or excluding petioles (64,838 observations and 7,558 species), and
2) a leaf disc excluding all major veins or petioles (645 observations and 403 species) [@Perez-Harguindeguy2013; @Kattge2020 [TRY version 5]].
Although many studies do not make clear what protocol was followed (146,315 observations and 13,101 species), these differences in the measurement methods may lead to large discrepancies in the estimates of LMA values because the major vein allocation (major vein, which includes first to second or third-order veins, volume per area) has been reported to be one of the main determinants for the variation in LMA [@John2017].

Since larger leaves tend to invest more of their mass into dense midribs and petiole for support [@Niinemets2006; @Niinemets2007], and thinner leaves have clearly visible and large-diameter veins with less uniform leaf structure (i.e., kite-type leaves; [@Grubb1986]), discrepancies in the estimates of whole-leaf LMA and leaf disc LMA might be greater for larger and thinner leaves.
If leaf disc LMA largely and consistently underestimates whole-leaf LMA, some calibrations may be required to combine or compare those different estimates [@Kraft2008; @Onoda2011].
If divergencies between the estimates based on the different methods are large and inconsistent (i.e., low *R^2^* values), this is difficult to calibrate and should inflate errors in the subsequent analyses.
To date, however, only a database from a single region, Panama Plant Traits Database [@Wright2010], is available in the TRY [@Kattge2020] that have both estimates of LMA from small leaf discs and whole leaves including petioles, limiting of our understanding of under what conditions estimates of LMA from leaf discs are reliable estimates of whole-leaf LMA.
